---
title: "Career Transition Tool"
output:
  flexdashboard::flex_dashboard:
    theme:
      version: 4
      primary: '#1f4181'
    source_code: https://github.com/bcgov/career_transition_proto
runtime: shiny
resource_files:
- processed_data/cip_noc_top5.csv
- processed_data/job_openings.csv
- processed_data/wages.csv
- processed_data/scaled_characteristics_noc.csv
- processed_data/teer_description.csv
- processed_data/unscaled_characteristics_noc.csv
---

```{r}
library(tidyverse)
library(vroom)
library(here)

#functions-----------------------
teer_table <- function(noc){
  teer_description%>%
    filter(teer==str_sub(noc,2,2))%>%
    select("{noc}":= description)%>%
    DT::datatable(rownames = FALSE, options = list(dom="t"))
}
make_plot <- function(tbbl){
  temp <- tbbl%>%
    group_by(category)%>% 
    mutate(all_category=fct_reorder(all_category, `Need for improvement`, .desc = TRUE))
 
  ggplot(temp, aes(all_category,
                  `Need for improvement`,
                  fill=category,
                  text=subcategory))+
    geom_col()+
    labs(y="Need for improvement",
         x="",
         fill="")+
    scale_y_continuous(limits = c(0, NA), labels = NULL)+
    scale_x_discrete(labels  = NULL) +
    scale_fill_brewer(palette = "Dark2")+
    theme_minimal()
}

make_plotly <- function(plt){ 
  plotly::ggplotly(plt, tooltip = "text")%>% 
    plotly::config(displayModeBar = F)
}
wage_plot <- function(){
  wages%>%
    filter(noc %in% c(input$noc, input$destination))%>%
    ggplot(mapping = aes(y=noc, 
                         x = median_wage,
                         text=paste("The typical hourly wage is", 
                                    scales::dollar(median_wage),
                                    "but can be as low as",
                                    scales::dollar(low_wage),
                                    "or as high as",
                                    scales::dollar(high_wage),
                                     "for <br>",
                                    str_sub(noc,8,-1)
                                    ))) +
    geom_pointrange(mapping = aes(xmin = low_wage, xmax = high_wage))+
    scale_x_continuous(labels=scales::dollar)+
    labs(y="", 
         x="Hourly Wage Rate")+
    theme_minimal()+ 
    theme(text = element_text(size = 12))
}
jo_plot <- function(num){
  job_openings%>%
    filter(noc %in% c(input$noc, input$destination))%>%
    group_by(noc)%>%
    summarize(sum_jo=sum(job_openings))%>%
    ggplot(mapping = aes(sum_jo, noc))+
    geom_col(alpha=.5)+
    theme_minimal()+
    labs(x="",
         y="")+
    scale_x_continuous(labels = scales::comma)
}


get_fos <- function(noc_code){
  cip_noc_top5%>%
  filter(noc == noc_code)%>%
  ungroup()%>%
  select(`Field of Study`)%>%
  pull()
}

fos_table <- function(){
  origin <- get_fos(input$noc)
  destination <- get_fos(input$destination)
  if(length(origin)==5 & length(destination)==5){
    fos_table <- tibble("{input$noc}":=origin, "{input$destination}":=destination)
  }else if(length(origin)==5){
    fos_table <- tibble("{input$noc}":=origin)
  }else if(length(destination)==5){
    fos_table <- tibble("{input$destination}":=destination)
  }else{
    fos_table <- tibble(sorry="no data")
  }
  fos_table%>%
    DT::datatable(rownames = FALSE, options=list(dom='t'))
}
reverse_legend_labels <- function(plotly_plot) {
  n_labels <- length(plotly_plot$x$data)
  plotly_plot$x$data[1:n_labels] <- plotly_plot$x$data[n_labels:1]
  plotly_plot
}



#load the data------------------------
teer_description <- vroom(here("processed_data","teer_description.csv"), delim =",")
cip_noc_top5 <- vroom(here("processed_data","cip_noc_top5.csv"))
wages <- vroom(here("processed_data","wages.csv"))
job_openings <- vroom(here("processed_data","job_openings.csv"))
first_year <- min(job_openings$year)
last_year <- max(job_openings$year)
scaled_char <- vroom(here("processed_data","scaled_characteristics_noc.csv"))%>%
  mutate(teer=as.numeric(str_sub(noc,2,2)), .after=noc)
categories <- unique(str_split_fixed(colnames(scaled_char)[-c(1:2)], pattern = ": ", n=2)[,1])
unscaled_char <- vroom(here("processed_data","unscaled_characteristics_noc.csv"))
```

## Sidebar {.sidebar}

-   Choosing an origin occupation in the first dropdown box populates the second dropdown box with ten closest destination occupations.
-   Closeness is based on 161 attributes from 4 categories:
    -   Abilities
    -   Knowledge
    -   Skills
    -   Work activities

```{r}
#inputs-------------------------
selectInput(
  "noc",
  "Choose an origin occupation:",
  choices = scaled_char$noc,
  selected = "41401: Economists and economic policy researchers and analysts",
  multiple = FALSE
)

checkboxInput("filter", "Only consider occupations with same or lower typical education and experience?", FALSE)

renderUI({#need the renderUI because the list of choices "nn()" is a reactive object
  selectInput('destination', 
              'Choose a destination occupation:', 
              choices = nn(), 
              multiple = FALSE
              )
})

selectInput(
  "category",
  "Choose what to plot top left:",
  choices = categories,
  selected = categories[1],
  multiple = FALSE
)


```

### Notes:

-   The top fields of study are not available for 11 occupations.

```{r}
#reactive elements----------------
tbbl <- reactive({
  if(input$filter==TRUE){
    scaled_char <- scaled_char%>%
      filter(teer>=as.numeric(str_sub(input$noc,2,2)))
  }
    scaled_char%>%
      select(-teer)%>%
      column_to_rownames(var="noc")
})

pca <- reactive({
  prcomp(tbbl())
})

first_pca <- reactive({ 
  pca()[["x"]][,1:5]%>% #keep only the first 5 principal components.
    as.data.frame()
})

q <- reactive({
  first_pca()[rownames(first_pca())==input$noc,] #the location of the chosen NOC in 5D space.
})

mat <- reactive({
  first_pca()[rownames(first_pca())!=input$noc,] #matrix of every other occupation
})

temp <- reactive({
  dbscan::kNN(mat(), k = 10, sort=TRUE,  query = q()) #10 nearest neighbors to query
})
nn <-  reactive({
  rownames(mat())[as.vector(temp()[["id"]])] #the names of the closest occupations
})
distance <- reactive({#gets the distance from the origin to the destination noc
  tibble(noc=nn(), distance=round(as.vector(temp()[["dist"]]), 2))%>%
    filter(noc==input$destination)%>%
    pull(distance)
})

compare <- reactive({
  unscaled_char[unscaled_char$noc %in% c(input$noc, input$destination),]%>%
  pivot_longer(cols = -noc)%>%
  separate(name, c("category", "subcategory"), sep=": ")%>%
  filter(category==input$category)%>%
  pivot_wider(names_from = noc, values_from = value)%>%
  slice_max(get(input$destination), n = 5, with_ties = FALSE)%>%
  pivot_longer(cols=-c("category","subcategory"), names_to= "noc", values_to = "value")
})

```

## Column

### **`r renderUI({paste0(input$destination, ", Distance = ", distance())})`**

```{r}
plotly::renderPlotly({
  plt <- ggplot(compare(), aes(value, fct_reorder(subcategory, value, max), fill=fct_rev(noc)))+
  geom_col(position = "dodge")+
  theme_minimal()+
  labs(x="",
       y="",
       fill="",
       title=paste0('<sup>', 
                    "Top 5 ", input$category, " for ",
                    str_sub(input$destination, 8),
                    '</sup>'),
       )+
   scale_x_continuous(limits=c(0,7))
  
  make_plotly(plt)%>%
  plotly::layout(
    title = list(x=-10),
    legend = list(
      orientation = "h",
      y=-.25
    )
  )%>%
  reverse_legend_labels()
})
```

### **Origin TEER**

```{r}
DT::renderDT({
  teer_table(input$noc)
})
```

### **Destination TEER**

```{r}
DT::renderDT({
  teer_table(input$destination)
})
```

## Column

### **Comparison of Hourly Wage Rates. (artificial data until available)**

```{r, fig.retina=3}
plotly::renderPlotly({
  make_plotly(wage_plot())
})
```

### **LMO Projected 10 year Job Openings (artifical data until release)**

```{r, fig.retina=3}
plotly::renderPlotly({
  make_plotly(jo_plot())
})
```

### **Top fields of study: % of workers with background**

```{r}
DT::renderDT({
  fos_table()
})

```
