---
title: "Career Transition Tool"
output: flexdashboard::flex_dashboard
runtime: shiny
resource_files:
- processed_data/cip_noc_top5.csv
- processed_data/job_openings.csv
- processed_data/wages.csv
- processed_data/scaled_characteristics_noc.csv
---

```{r}
library(tidyverse)
library(vroom)
library(here)
#functions-----------------------
make_plot <- function(tbbl){
  temp <- tbbl%>%
    group_by(category)%>% 
    mutate(all_category=fct_reorder(all_category, `Need for improvement`, .desc = TRUE))
 
  ggplot(temp, aes(all_category,
                  `Need for improvement`,
                  fill=category,
                  text=subcategory))+
    geom_col()+
    labs(y="Need for improvement",
         x="",
         fill="")+
    scale_y_continuous(limits = c(-2, 4), labels = NULL)+
    scale_x_discrete(labels  = NULL) +
    scale_fill_brewer(palette = "Dark2")+
    theme_minimal()
}

make_plotly <- function(plt){ 
  plotly::ggplotly(plt, tooltip = "text")%>% 
    plotly::config(displayModeBar = F)
}
wage_plot <- function(num){
  wages%>%
    filter(noc %in% c(input$noc, difference()$noc[num]))%>%
    ggplot(mapping = aes(y=noc, x = median_wage)) +
    geom_pointrange(mapping = aes(xmin = low_wage, xmax = high_wage))+
    scale_x_continuous(labels=scales::dollar)+
    labs(y="", 
         x="Hourly Wage Rate")+
    theme_minimal()+ 
    theme(text = element_text(size = 12))
}
jo_plot <- function(num){
  job_openings%>%
    filter(noc %in% c(input$noc, difference()$noc[num]))%>%
    ggplot(mapping = aes(year, job_openings, colour=noc)) +
    geom_line()+
    theme_minimal()+
    theme(legend.position = "bottom")+
    guides(colour=guide_legend(nrow=2,byrow=TRUE))+
    labs(y="",
         x="",
         colour="")+ 
    theme(text = element_text(size = 17))
}

fos_table <- function(num){
  origin <- cip_noc_top5%>%
  filter(noc ==input$noc)%>%
  ungroup()%>%
  select(`Field of Study`)%>%
  pull()

destination <- cip_noc_top5%>%
  filter(noc == difference()$noc[num])%>%
  ungroup()%>%
  select(`Field of Study`)%>%
  pull()

tibble("{input$noc}":=origin,
       "{difference()$noc[num]}":=destination)%>%
  DT::datatable(rownames = FALSE, options=list(dom='t'))
}

#load the data, do the PCA
cip_noc_top5 <- read_csv(here("processed_data","cip_noc_top5.csv"))
wages <- vroom(here("processed_data","wages.csv"))
job_openings <- vroom(here("processed_data","job_openings.csv"))
tbbl <- vroom(here("processed_data","scaled_characteristics_noc.csv"))%>%
  column_to_rownames(var="noc")
pca <- prcomp(tbbl)
first_pca <- pca[["x"]][,1:5]%>% #keep only the first 5 principal components.
    as.data.frame()
```

Sidebar {.sidebar}
=====================================

* When you choose an occupation from the list below, this app finds the 10 closest occupations.
* Closeness is based on 161 attributes from 4 categories:
    - Abilities
    - Knowledge
    - Skills
    - Work activities 
* The ten closest occupations are arranged in tabs across the top of the page.
* The left panel show the attributes most in need of improvement in order to switch occupations: a blank plot indicate an exact match to chosen occupation.
* In the right panels there is a comparison between the origin and destination occupations in terms of wages, job openings and typical educational background.

```{r}
selectInput(
  "noc",
  "Choose an occupation:",
  rownames(tbbl),
  selected = "41401: Economists and economic policy researchers and analysts",
  multiple = FALSE
)
```

```{r}
#reactive elements----------------
q <- reactive({
  first_pca[rownames(first_pca)==input$noc,] #the location of the chosen NOC in 10D space.
})
temp <- reactive({
  dbscan::kNN(first_pca, k = 11, sort=TRUE,  query = q()) #11 nearest neighbors to query (own NOC included)
})
nn <-  reactive({
  rownames(first_pca)[as.vector(temp()[["id"]])] #the names of the closest occupations
})
chosen_noc <- reactive({
  tbbl[rownames(tbbl)==input$noc,] #the skill profile of the chosen NOC (35D)
})
repeated_chosen <- reactive({
  chosen_noc()[rep(1, 11), ] #replicate the skill profile of the chosen NOC
})
closest <- reactive({
  tbbl[rownames(tbbl) %in% nn(), ] #the nearest neighbour NOCs to the chosen NOC
})
wrong_order_difference <- reactive({
  (closest()-repeated_chosen())%>% #the difference in skills
  mutate(across(everything(), \(x) round(x, 2)))%>%
  rownames_to_column(var="noc")%>%
  pivot_longer(cols=-noc, names_to = "all_category", values_to = "Need for improvement")%>%
  mutate(temp=all_category)%>%
  separate(temp, into = c("category","subcategory"), sep=": ")%>%
  filter(noc!=input$noc)%>%
  group_by(noc, category)%>%
  slice_max(order_by = `Need for improvement`, n = 5, with_ties=FALSE)%>%
  group_by(noc)%>%
  nest()%>%
  mutate(plot=map(data, make_plot),
         plotly=map(plot, make_plotly)
         )
})
#order the dataframe correctly------------
difference <- reactive({
  wrong_order_difference()[match(nn(), wrong_order_difference()$noc), ]%>%
  na.omit()
})
#the pages--------------
```


`r renderUI({paste0(word(str_split_1(difference()$noc[[1]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### **`r renderUI({str_split_1(difference()$noc[[1]], ":")[2]})`**
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[1]]
}) 
```

### TEER

   
Column 
-------------------------------------
   
### Comparison of Hourly Wage Rates.

```{r, fig.retina=3}
plotly::renderPlotly({
  plotly::ggplotly(wage_plot(1))%>% 
    plotly::config(displayModeBar = F)
})
```   
 
### LMO Projected Job Openings for years 2024 to 2033.
    
```{r, fig.retina=3}
renderPlot({
  jo_plot(1)
})
```

### Top fields of study: % of workers with background

```{r}
DT::renderDT({
  fos_table(1)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[2]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[2]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[2]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(2)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(2)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(2)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[3]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[3]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[3]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(3)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(3)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(3)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[4]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[4]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[4]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(4)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(4)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(4)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[5]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[5]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[5]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(5)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(5)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(5)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[6]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[6]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[6]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(6)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(6)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(6)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[7]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[7]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[7]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(7)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(7)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(7)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[8]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[8]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[8]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(8)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(8)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(8)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[9]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[9]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[9]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(9)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(9)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(9)
})

```

`r renderUI({paste0(word(str_split_1(difference()$noc[[10]], ":")[2],2),"...")})`
=====================================  

Column 
-------------------------------------
    
### `r renderUI({str_split_1(difference()$noc[[10]], ":")[2]})`
    

```{r}
plotly::renderPlotly({
  difference()[["plotly"]][[10]]
}) 
```
   
Column 
-------------------------------------
   
### Wage comparison

```{r}
renderPlot({
  wage_plot(10)
})
```   
 
### Job Opening comparison
    
```{r}
renderPlot({
  jo_plot(10)
})
```

### Field of Study comparison

```{r}
DT::renderDT({
  fos_table(10)
})

```



